StreamReduce: Processing Data Streams\\and Avoiding Work

### 2012-05-10 paper/paper.tex
Editing pass. The end is near

\section{Introduction}
\label{sec:intro}
One of the fundamental problems MapReduce was created to solve was processing
the growing amount of data in the world.
MapReduce lends itself naturally to top-k-style computations,
data sampling tasks, and other cases where perfect-fidelity results are not required.
Now with the open source Apache Hadoop
and Pig~\cite{Olston:2008:PLN:1376616.1376726}, it is easy to process and query
the terabytes of data an organization may collect.
%The widespread use of these tools to solve problems they weren't designed for motivates
%us to explore how we can save time, resources, and prevent faults in the execution of
%such MapReduce tasks.

@@

The \emph{Data Stream Model}~\cite{Babcock:2002:MID:543613.543615} is a refinement of the Chronicle
Data Model. It has the following properties: stream items may arrive out of order,
streams are potentially unbounded in size, elements in the stream are discared after they are
processed. This is the data model that StreamReduce uses.

@@

Dynamic Knobs~\cite{Hoffmann:2011:DKR:1950365.1950390} describes a mechanism to explore the
accuracy-performance tradeoff space to adjust dynamically adjust program tuning parameters
to changes in the execution environment. This allows programs to do things such as
utilize more resources if they become available, use less power if power rates spike,
or use less accurate computations in the face of an impending time constraint.

@@

Collectors also expose an API endpoint for accessing the results of a job. This endpoint
can be used by another spout to chain StreamReduce jobs together.
\subsection{Master}
\label{sec:master}
There is one master node per StreamReduce network. It has a function similar to the jobtracker in
MapReduce. New jobs are scheduled with the master and it initializes other nodes in the
cluster with job state. For each job, the master schedules roles among nodes in a round robin
manner to minimize the total number of outstanding roles on any one node.

@@

StreamReduce sends the source of the Jobfile to every node participating in the job and uses eval to
obtain a local instance of the Jobfile class. An example jobfile can be found at~\cite{github}.

The use of eval to instantiate Jobfile objects and the implementation language's (ruby)
ability to monkeypatch classes means that jobs can be modified on the fly just by sending
an updated Jobfile to job participants. For example, a job can be made lower fidelity in
the face of an impending time constraint, be made higher fidelity if more workers are added to
the cluster, or gracefully kill all fetchers which allows the job to end after all of the
various job queues are empty.
\subsection{Fault Tolerance}
StreamReduce can sustain the failure of any node other than the master. Because the master is responsible
for shuffling data between fetchers, workers, and collectors, a job cannot make progress without
the master.

@@

Workers are initialized with the~3 computation routines shown in Table~\ref{table:workerTasks}.

@@

\subsection{Precision and Recall}
In order for task killing to be an effective strategy for a top-k computation, the
distribution of words in the corpus must be able to sustain data loss. This is the case
if the distribution is long-tailed, but is not true if the distribution is uniform.

@@

Poor recall statistics are acceptable because I am running a top-k job that focuses
on good precision in the top~100. Because I use a weighted $F$-score, the poor recall's impact
on measured accuracy is mitigated.

@@

The average reduction in $F_{0.5}$-score caused by reducing the work level form 0.8 to 0.4 is
$0.231$ with a standard deviation of $0.025$. I performed four additional experiments with the
work level set to $0.8$ and $0.4$ with the task kill rate set to $0.1$ and $0.7$. These experiments were
not included in generating Figure~\ref{fig:fscore}. The average reduction in $F_{0.5}$-score between
work level $0.8$ and $0.4$ in these experiments is $0.230$ with a standard deviation of
$0.056$. These numbers show that a $0.4$ reduction in the work level leads to approximately a
$0.23$ reduction in $F_{0.5}$-score.

@@

In order to mitigate the effects of this problem, I batched individual requests into groups of~5. This
lowered the number of connections I was initiating but it also limited the parallelism of the cluster
and removed workers from the pool for longer periods because they had more data to process at a time.
A killed task on the worker no longer freed it up to receive a new task immediately.
Keeping workers removed from the pool caused the queues on the master to grow longer as the workers
could no longer keep up with the rate the fetcher was producing data.

@@

In order to address this problem, the master needs to keep a persistent channel open between
it and all active nodes. Communication channels should be established when nodes come online.
Possible solutions include writing to a socket or using a higher abstraction like AMPQ\@.
AMPQ is appealing because it also relieves the master from handling queues.

>>>>

\section{Introduction}
\label{sec:intro}
One of the fundamental problems MapReduce was created to solve is processing
the growing amount of data in the world.
MapReduce lends itself naturally to top-k-style computations,
data sampling tasks, and other cases where perfect-fidelity results are not required.
Now with the open source Apache Hadoop
and Pig~\cite{Olston:2008:PLN:1376616.1376726}, it is easy to process and query
the terabytes of data an organization may collect.
%The widespread use of these tools to solve problems they weren't designed for motivates
%us to explore how we can save time, resources, and prevent faults in the execution of
%such MapReduce tasks.

@@

The \emph{Data Stream Model}~\cite{Babcock:2002:MID:543613.543615} is a refinement of the Chronicle
Data Model. It has the following properties: stream items may arrive out of order,
streams are potentially unbounded in size, and elements in the stream are discared after they are
processed. This is the data model that StreamReduce uses.

@@

Dynamic Knobs~\cite{Hoffmann:2011:DKR:1950365.1950390} describes a mechanism to explore the
accuracy-performance tradeoff space, dynamically adjusting program tuning parameters in response
to changes in the execution environment. This allows programs to do things such as
utilize more resources if they become available, use less power if power rates spike,
or use less accurate computations in the face of an impending time constraint.

@@

Collectors also expose an API endpoint for accessing the results of a job. This endpoint
can be used by another spout to chain StreamReduce jobs together.
\subsection{Master}
\label{sec:master}
There is one master node per StreamReduce cluster. It has a function similar to the jobtracker in
MapReduce. New jobs are scheduled with the master and it initializes other nodes in the
cluster with job state. For each job, the master schedules roles among nodes in a round robin
manner to minimize the total number of outstanding roles on any one node.

@@

StreamReduce sends the source of the Jobfile to every node participating in the job and uses
\texttt{eval} to
obtain a local instance of the Jobfile class. An example jobfile can be found at~\cite{github}.

The use of \texttt{eval} to instantiate Jobfile objects and the implementation language's (ruby)
ability to monkeypatch classes means that jobs can be modified on the fly just by sending
an updated Jobfile to job participants. For example, a job can be made lower fidelity in
the face of an impending time constraint, be made higher fidelity if more workers are added to
the cluster, or gracefully kill all fetchers which allows the job to end after all of the
various job queues are empty.
\subsection{Fault Tolerance}
StreamReduce can sustain the failure of any node other than the master. Because the master is responsible
for shuffling data between fetchers, workers, and collectors, a job cannot make progress without
the master.

@@

Workers are initialized with the three computation routines shown in Table~\ref{table:workerTasks}.

@@

\subsection{Precision and Recall}
For task killing to be an effective strategy for a top-k computation, the
distribution of words in the corpus must be able to sustain data loss. This is the case
if the distribution is long-tailed, but is not true if the distribution is uniform.

@@

Poor recall statistics are acceptable because I am running a top-k job that focuses
on good precision in the top~100. Because I use a weighted $F$-score, the poor recall's impact
on measured accuracy is lessened.

@@

The average reduction in $F_{0.5}$-score caused by reducing the work level form 0.8 to 0.4 is
$0.231$ with a standard deviation of $0.025$. I performed four additional experiments with the
work level set to $\{0.8,0.4\}$ and with the task kill rate set to $\{0.1,0.7\}$. These experiments were
not included in generating Figure~\ref{fig:fscore}. The average reduction in $F_{0.5}$-score between
work level $0.8$ and $0.4$ in these experiments is $0.230$ with a standard deviation of
$0.056$. These numbers show that a $0.4$ reduction in the work level leads to approximately a
$0.23$ reduction in $F_{0.5}$-score.

@@

To mitigate the effects of this problem, I batched individual requests into groups of~5. This
lowered the number of connections I was initiating but it also limited the parallelism of the cluster
and removed workers from the pool for longer periods because they had more data to process at a time.
A killed task on the worker no longer freed it up to receive a new task immediately.
Keeping workers removed from the pool caused the queues on the master to grow longer as the workers
could no longer keep up with the rate the fetcher was producing data.

@@

To address this problem, the master needs to keep a persistent channel open between
it and all active nodes. Communication channels should be established when nodes come online.
Possible solutions include writing to a socket or using a higher abstraction like AMPQ\@.
AMPQ is appealing because it also relieves the master from handling queues.
